---
title: "RTconflict"
author: "Thibault Gajdos, Mathieu Servant, Thierry Hasbroucq, Karen Davranche"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, SETTINGS-knitr, include=FALSE}
stopifnot(require(knitr))
options(width = 90)
opts_chunk$set(
  comment = NA,
  ##message = FALSE,
  ##warning = FALSE,
  fig.align = "center"
  )
```

## Contents {#contents}

  1. [Overview](#overview)
  2. [Data](#data)
  3. [ELF](#ELF)
  4. [Delta plots](#delta)

## Overview {#overview}

This package provides tools to analyse response time (RT) distributions for conflict tasks in R. Specifically, it allows to compute: 
- the Error Location Function and the Error Location Index proposed in Servant, M., Gajdos, T., & Davranche, K. (2018). ELF: A new measure of response capture. *Psychonomic Bulletin & Review*, 1-9.

- Lorenz-inhibition index proposed in Gajdos, T., Servant, M., Hasbroucq, T., & Davranche, K. (2019, March 21). Assessing Inhibitory Control from Reaction Time Distributions: The Lorenz Inhibition Index. https://doi.org/10.31234/osf.io/7f5n8


To install, use the following command: 

`devtools::install_github("thibault-gajdos/RTconflict")`

## Data {#data}

We use simulated data for a conflict task. 
```{r  echo = FALSE, results = "hide", warning = FALSE, message = FALSE}
library(RTconflict)
library(tidyverse)
```

There are 20 subjects (from 1 to 20), performing a conflict task with 200 congruent and 200 incongruent trials, in two distinct conditions (A and B). Therefore, there is 800 trials by subject.


```{r}
data.conflict %>%
    group_by(compatible, condition) %>%
    summarise(acc = mean(acc), rt = mean(rt))
```

## Error Location Function {#ELF}

We start with a single subject (subject 1). Error Location Functions are particularly meaningfull for incongruent trials. With thus restrict attention to these trials.

First, we compute ELF in each condition.

```{r}
d1A <- data.conflict %>% filter(sujet == 1, condition == 'A', compatible == 'i')
p.1A <- c(1:nrow(d1A))/nrow(d1A)
elf.1A <- elf(d1A$rt, d1A$acc)
df1A <- data.frame(elf = elf.1A, p = p.1A)  %>%
    mutate(condition = 'A')

d1B <- data.conflict %>% filter(sujet == 1, condition == 'B', compatible == 'i')
elf.1B <- elf(d1B$rt, d1B$acc)
p.1B <- c(1:nrow(d1B))/nrow(d1B)
df1B <- data.frame(elf = elf.1B, p = p.1B) %>%
    mutate(condition = 'B')

df1 <- rbind(df1A, df1B)
```

We now plot ELF.

```{r}
plot.elf1 <- ggplot(data = df1, aes(x = p, y = elf, group = condition, colour = condition)) +
    geom_line() +
    xlab('p') +
    ylab('% of total errors') + 
    ggtitle('Error Location Function')
plot.elf1
```

The Error Location Index (ELI) is the area under ELF. 

```{r}
eli(elf.1A)
eli(elf.1B)
```

So, as suggested by the ELF plots, response capture seems larger in condition B than in condition A. Now, we will turn to aggregate ELF accross subects. 

```{r}
dA <- data.conflict %>% filter(compatible == 'i', condition == 'A')
elf.A <- elf.g(dA$rt, dA$acc, dA$sujet) 

dB <- data.conflict %>% filter(compatible == 'i', condition == 'B')
elf.B <- elf.g(dB$rt, dB$acc, dB$sujet)
```

We can now plot these aggregated ELF functions.

```{r}
d <- data.frame(p = c(0:100,0:100), elf = c(elf.A, elf.B), condition = c(rep('A',101),rep('B',101)))

plot.elf <- ggplot(data = d, aes(x = p, y = elf, group = condition, colour = condition)) +
    geom_line() +
    xlab('p') +
    ylab('% of total errors') + 
    ggtitle('Aggregated Error Location Function')
plot.elf
```

## Delta plots {#delta}

To study inhibition, we focus on correct trials. We again start by focusing on one subject, and this time consider only one condition (A). We first compute the absolute Lorenz-delta function. 


```{r}
data1A <- data.conflict %>% filter(sujet == 1, condition == 'A', acc == 1)
LD.1A <- lorenz(data1A$rt, data1A$compatible)
glimpse(LD.1A)
```
LD.1A contains two parts: 

-  inhib which is the Lorenz-inhibition index

- q which contains:

	- p: quantiles

	- qc: quantile function for compatible trials

	- qi: quantile function for incompatible trials
 
	- delta = (qi-qc)/2

	- m = (qc+qi)/2

	- rdelta = delta/m
  
	- c = cumulative function of rdelta 
  
Let's plot the Lorenz-delta plot.

```{r}
plot.1A <- ggplot(data = LD.1A$q, aes(x = p, y = c)) +
    geom_line(size = 1) +
    xlab('p') +
    ylab('% R(p)') +
    ggtitle('Lorenz-delta plot') 
plot.1A
```
	
Now, let compute an aggregate  Lorenz-delta plot accross subjects for each condition. First we compute  Lorenz-delta plots for each subject and each condition. Note that it might be the case that different subjects or conditions have different numbers of quantiles. To solve this, we constrain the number of quantiles by grouping quantiles by 1%.

```{r}
x  <- NULL 
for (s in unique(data.conflict$sujet)){
    for (cond in unique(data.conflict$condition)){ 
    d  <- data.conflict %>% filter(sujet == s, condition == cond, acc == 1) 
    xx  <- lorenz(d$rt, d$compatible)$q %>%
                                   mutate(p  = round(p, digits = 2)) %>%
                                   group_by(p) %>% 
                                   summarise(c = mean(c)) %>%
                                   mutate(sujet = s, condition = cond)
    x  <- rbind(x, xx)
    }
}
```

Now we aggregate these Lorenz-Delta plots together

```{r}
x.ag  <- x %>%
    group_by(p, condition) %>%
    summarise(c = mean(c))
```

and plot them.

```{r}
plot  <- ggplot(data = x.ag, aes(x = p, y = c, group = condition, colour = condition)) +
    geom_line(size = 1) +
    xlab('p') +
    ylab('% R(p)') +
    ggtitle('aggregated  Lorenz-delta plot')
plot
```

`data.inhib`  computes delta plots and Lorenz-inhibition index for each subject and each condition.

```{r}
data.inhib <- data.conflict %>% filter(acc == 1)
inhib  <- inhib.delta(rt = data.inhib$rt, comp = data.inhib$compatible, sujet = data.inhib$sujet, cond = data.inhib$condition)
```

The function generates three data frame:
- i: Lorenz-inhibition index by subject and condition
- delta.slope: slope of the last segment of the delta plot by subject and condition. This is just for information, as we strongly advice against the use of this measure.


```{r}
glimpse(inhib)
```

The mean of the Lorenz-inhibition index is 
```{r}
inhib$i %>%  group_by(cond) %>%
    summarise(index = mean(index)) 
```

