---
title: "RTconflict"
author: "Thibault Gajdos, Mathieu Servant, Thierry Hasbroucq, Karen Davranche"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, SETTINGS-knitr, include=FALSE}
stopifnot(require(knitr))
options(width = 90)
opts_chunk$set(
  comment = NA,
  ##message = FALSE,
  ##warning = FALSE,
  fig.align = "center"
  )
```

## Contents {#contents}

  1. [Overview](#overview)
  2. [Data](#data)
  3. [ELF](#ELF)
  4. [Interference Distribution Index](#delta)

## Overview {#overview}

This package provides tools to analyse response time (RT) distributions for conflict tasks in R. Specifically, it allows to compute: 

- the Error Location Function and the Error Location Index proposed in Servant, M., Gajdos, T., & Davranche, K. (2018). ELF: A new measure of response capture. *Psychonomic Bulletin & Review*, 1-9.

- Interference Distribution Index proposed in Gajdos, T., Servant, M., Hasbroucq, T., & Davranche, K. (2019, december 20). A measure of the Interference Effect Distribution. https://doi.org/10.31234/osf.io/7f5n8


To install, use the following command: 

`devtools::install_github("thibault-gajdos/RTconflict")`

## Data {#data}

We use simulated data for a conflict task. 
```{r  echo = FALSE, results = "hide", warning = FALSE, message = FALSE}
library(RTconflict)
library(tidyverse)
```

There are 20 subjects (from 1 to 20), performing a conflict task with 200 congruent and 200 incongruent trials, in two distinct conditions (A and B). Therefore, there is 800 trials by subject.


```{r}
data.conflict %>%
    group_by(compatible, condition) %>%
    summarise(acc = mean(acc), rt = mean(rt))
```

## Error Location Function {#ELF}

We start with a single subject (subject 1). Error Location Functions are particularly meaningfull for incongruent trials. With thus restrict attention to these trials.

First, we compute ELF in each condition.

```{r}
d1A <- data.conflict %>% filter(sujet == 1, condition == 'A', compatible == 'i')
p.1A <- c(1:nrow(d1A))/nrow(d1A)
elf.1A <- elf(d1A$rt, d1A$acc)
df1A <- data.frame(elf = elf.1A, p = p.1A)  %>%
    mutate(condition = 'A')

d1B <- data.conflict %>% filter(sujet == 1, condition == 'B', compatible == 'i')
elf.1B <- elf(d1B$rt, d1B$acc)
p.1B <- c(1:nrow(d1B))/nrow(d1B)
df1B <- data.frame(elf = elf.1B, p = p.1B) %>%
    mutate(condition = 'B')

df1 <- rbind(df1A, df1B)
```

We now plot ELF.

```{r}
plot.elf1 <- ggplot(data = df1, aes(x = p, y = elf, group = condition, colour = condition)) +
    geom_line() +
    xlab('p') +
    ylab('% of total errors') + 
    ggtitle('Error Location Function')
plot.elf1
```

The Error Location Index (ELI) is the area under ELF. 

```{r}
eli(elf.1A)
eli(elf.1B)
```

So, as suggested by the ELF plots, response capture seems larger in condition B than in condition A. Now, we will turn to aggregate ELF accross subects. 

```{r}
dA <- data.conflict %>% filter(compatible == 'i', condition == 'A')
elf.A <- elf.g(dA$rt, dA$acc, dA$sujet) 

dB <- data.conflict %>% filter(compatible == 'i', condition == 'B')
elf.B <- elf.g(dB$rt, dB$acc, dB$sujet)
```

We can now plot these aggregated ELF functions.

```{r}
d <- data.frame(p = c(0:100,0:100), elf = c(elf.A, elf.B), condition = c(rep('A',101),rep('B',101)))

plot.elf <- ggplot(data = d, aes(x = p, y = elf, group = condition, colour = condition)) +
    geom_line() +
    xlab('p') +
    ylab('% of total errors') + 
    ggtitle('Aggregated Error Location Function')
plot.elf
```

## Interference Effect Distribution {#delta}

To study inhibition, we focus on correct trials. We again start by focusing on one subject, and this time consider only one condition (A). We  compute the Interference Distribution Index. 


```{r}
data1A <- data.conflict %>% filter(sujet == 1, condition == 'A', acc == 1)
LD.1A <- lorenz(data1A$rt, data1A$compatible)
glimpse(LD.1A)
```
As an option, one can specify the quantile method by setting "type" to a value between 1 and 9, 7 be the default. See the documentation of the quantile function in R for further details. 

LD.1A contains two parts: 

-  inhib which is the Lorenz-inhibition index

- q which contains:

	- p: quantiles

	- qc: quantile function for compatible trials

	- qi: quantile function for incompatible trials
 
	- delta = (qi-qc)/2

	- m = (qc+qi)/2
  
	- c = cumulative function of delta 
  
Let's plot the Lorenz-interference plot.

```{r}
plot.1A <- ggplot(data = LD.1A$q, aes(x = p, y = c)) +
    geom_line(size = 1) +
    xlab('p') +
    ylab('% D(p)') +
    ggtitle('Lorenz-inhibition plot') 
plot.1A
```
	
Now, let compute an aggregate  Lorenz-inhibition plot accross subjects for each condition. First we compute  Lorenz-inhibition plots for each subject and each condition. Note that it might be the case that different subjects or conditions have different numbers of quantiles. To solve this, we constrain the number of quantiles by grouping quantiles by 1%.

```{r}
x  <- NULL 
for (s in unique(data.conflict$sujet)){
    for (cond in unique(data.conflict$condition)){ 
    d  <- data.conflict %>% filter(sujet == s, condition == cond, acc == 1) 
    xx  <- lorenz(d$rt, d$compatible)$q %>%
                                   mutate(p  = round(p, digits = 2)) %>%
                                   group_by(p) %>% 
                                   summarise(c = mean(c)) %>%
                                   mutate(sujet = s, condition = cond)
    x  <- rbind(x, xx)
    }
}
```

Now we aggregate these Lorenz-Delta plots together

```{r}
x.ag  <- x %>%
    group_by(p, condition) %>%
    summarise(c = mean(c))
```

and plot them.

```{r}
plot  <- ggplot(data = x.ag, aes(x = p, y = c, group = condition, colour = condition)) +
    geom_line(size = 1) +
    xlab('p') +
    ylab('% D(p)') +
    ggtitle('aggregated  Lorenz-inhibition plot')
plot
```

`data.inhib`  computes delta plots and the Inhibition Distribution Index for each subject and each condition.

```{r}
data.inhib <- data.conflict %>% filter(acc == 1)
inhib  <- inhib.delta(rt = data.inhib$rt, comp = data.inhib$compatible, sujet = data.inhib$sujet, cond = data.inhib$condition)
```

The function generates three data frame:
- i: Inhibition Distribution Index index by subject and condition
- delta.slope: slope of the last segment of the delta plot by subject and condition. This is just for information, as we strongly advice against the use of this measure.


```{r}
glimpse(inhib)
```

The mean of the Inhibition Distribution Index index is 
```{r}
inhib$i %>%  group_by(cond) %>%
    summarise(index = mean(index)) 
```

